### 2020-10-06 HW-AlgebraicTopology2020/algtop2/algtop2.tex
Restructure to gather quick notes

\newcommand{\hwNumber}{3 [WIP]}

@@

\newcommand{\vset}[1]{\Bqty{\ \begin{gathered}
	#1
\end{gathered}\ }}
\newcommand{\Zpoly}[1]{\mbb{Z}[x]\big/x^{#1}}
\newcommand{\colim}{\mop{colim}}

@@

	\begin{enumerate}
	\item \textbf{Example of limit in \Cat{Vect}:}
	\begin{equation}
	\begin{tikzcd}[row sep=-3.5ex,column sep=8ex]
		& \vset{0 \\ \mbb{R}} \\
		I = \vset{\bullet \\ \star}
			\arrow[ur,mapsto,%
				"F"]
			\arrow[dr,mapsto,swap,%
				"\Delta(V)" near end]
		& \\
		& \vset{V \\ V}
	\end{tikzcd}
		\ \supset\ \Cat{Vect}\colon \text{
			$\mbb{R}$ vector space
		},
	\end{equation}

	\begin{enumerate}
	\item By definition, we have:
	\begin{equation}
	\begin{tikzcd}[row sep=8.5ex,column sep=5ex]
		& \Delta(\lim F)
			\arrow[d,Rightarrow,"\sigma"] \\
		\Delta(V)
			\arrow[r,Rightarrow,"\eta"]
			\arrow[ur,Rightarrow,dashed,%
				"\exists!\,\Delta(g)"] &
		F
	\end{tikzcd}
	\label{eq:def_limit}
	\end{equation}
	Where $g\colon V\to \lim F$. 
	More intuitively, for the above $F\colon I\to\Cat{Vect}$, this translates to the following diagram in \Cat{Vect}:
	\begin{equation}
	\begin{tikzcd}[row sep=1.5ex,column sep=8ex]
		&& 0 \\
		V
			\arrow[r,dashed,
				"\exists!\, g"]
			\arrow[urr,bend left=25,%
				"\eta_\bullet"]
			\arrow[drr,bend right=25,swap,%
				"\eta_\star"] &
		\lim F
			\arrow[ur,%
				"\sigma_\bullet"]
			\arrow[dr,swap,%
				"\sigma_\star"] & \\
		&& \mbb{R}
	\end{tikzcd}
	\label{eq:lim_in_vect}
	\end{equation}
	\end{enumerate}

	Now we verify that $\lim F = \mbb{R}$, along with the following choice of $\sigma$:
	\begin{equation}
		\sigma_\bullet = 0,\quad
		\sigma_\star = \idty_\mbb{R}
	\end{equation}
	In fact, for the above diagram \eqref{eq:lim_in_vect} to be commutative, we must have $g = \eta_\star$. Note that such $g$ is unique once $\sigma$ is chosen; for our choice of $\sigma$, if $g \ne \eta_\star$, then the diagram \textit{cannot} commute. Hence $\lim F = \mbb{R}$, along with the above choice of $
		\sigma\colon \Delta(\mbb{R})\Rightarrow F
	$. In other words, we have:
	\begin{equation}
	\begin{tikzcd}[row sep=1ex,column sep=2.5em]
		& \Bqty\big{\ \bullet,\ \star\ }
			\arrow[ddl,"\Delta(\mbb{R})",swap]
			\arrow[ddr,"F"]
		& \\
		& \xRightarrow{\quad\sigma\quad} & \\
		\mbb{R} & & \Bqty\big{\ 0,\ \mbb{R}\ }
	\end{tikzcd}
	\end{equation}
	\vspace{-2.25\baselineskip}

	\qed

	\item[(b)(c)] From diagram \eqref{eq:def_limit} and discussions in (a), we know that: 
	\begin{equation}
		\exists\,!\ \tau = \Delta(g)\colon\ %
		\Delta(V)\Longrightarrow \Delta(\mbb{R})
	\end{equation}
	Here $g\colon V\to\mbb{R}$ is fixed uniquely once $\sigma$ is fixed. However, $\sigma$ may vary up to isomorphism; therefore, a generic choice of $\sigma$ is given by:
	\begin{equation}
		\sigma_\bullet = 0,\quad
		\sigma_\star = k\,\idty_\mbb{R},\quad
		k\in\mbb{R}
	\end{equation}

	For such $g$, by the same arguments in (a), we have:
	\begin{equation}
		\exists\,!\,\ %
			g = \frac{1}{k}\,\eta_\star\,,\ %
			\tau
			= \Delta(g)
			= \frac{1}{k}\,\Delta(\eta_\star),\quad
		\textsl{s.\,t.\quad
			\eqref{eq:def_limit},
			\eqref{eq:lim_in_vect} commutes}
	\end{equation}
	Note that $k\in\mbb{R}$, for every $k$ there is a different $g$ and $\tau$; hence there are $
		\norm{\{k\}} = \norm{\mbb{R}}
	$ many choices of $\tau$ to make the diagram commute. In particular, for $k = 1$ we recover $
		\tau = \Delta(\eta_\star)
	$. \qedfull

	\item \textbf{Limit and colimit of polynomial ring:}

	By definition,
	\begin{equation}
	\begin{tikzcd}[row sep=7ex,column sep=8ex,baseline=-.8ex]
		&&[-4ex] Q
			\arrow[ddll]
			\arrow[ddl,"\eta_{n+1}" near end,swap]
			\arrow[ddr,"\eta_{n}" near end]
			\arrow[ddrr]
			\arrow[ddrrr]
			\arrow[d,dotted,"\exists! f" description]
		&[-4ex] && \\
		&& \lim F
			\arrow[dl,"\pi_{n+1}"]
			\arrow[dr,"\pi_{n}",swap]
		&&& \\
		\cdots
			\arrow[ddrr]
			\arrow[r,"p_{n+1}"] &
		\Zpoly{n+1}
			\arrow[dr,"\sigma_{n+1}"]
			\arrow[ddr,
				"\tau_{n+1}\!" near start,swap]
			\arrow[rr,"p_n"] 
		&&
		\Zpoly{n}
			\arrow[dl,"\sigma_{n}",swap]
			\arrow[ddl,
				"\tau_{n}" near start]
			\arrow[r,"p_{n-1}"] &
		\cdots
			\arrow[ddll]
			\arrow[r,"p_1"] &
		\Zpoly{1} = \mbb{Z}
			\arrow[ddlll,
				"\tau_1" near start] \\
		&& \colim F
			\arrow[d,dotted,"\exists! g" description]
		&&& \\
		&& R
		&&&
	\end{tikzcd}
	\label{eq:polynomial_ring_tower}
	\end{equation}
	Here $p_n\colon\Zpoly{n+1}\to\Zpoly{n}$ is the natural projection.

	Intuitively, if such $\lim F$ exists, it shall be the ``smallest'' object that ``contains'' $\Zpoly{n}$ when $n\to\infty$. Note that $\Zpoly{n}$ is naturally a $\mbb{Z}^n$ vector space:
	\begin{equation}
		\Zpoly{n}\ \ni\ \sum_{m=0}^{n-1} a_m x^m
		\ \sim\ \pqty{
			a_0, a_1, \cdots, a_{n-1}
		}\ \in\ \mbb{Z}^n
	\end{equation}
	While $n\to\infty$, this gives an $\infty$--tuple which corresponds to the \textit{formal power series}\footnote{
		See \wikiref{https://en.wikipedia.org/wiki/Formal\_power\_series}{Formal power series}. This is in fact the \textit{adic completion} of $\mbb{Z}[x]$. I would like to thank \mbox{\textit{刘逸华}} \& \textit{谢贤进} for this hint. 
	}: 
	\begin{equation}
		\mbb{Z}[[x]] \ \ni\ \sum_{m=0}^\infty
			a_m x^m
	\end{equation}

	The difference between $\mbb{Z}[x]$ and $\mbb{Z}[[x]]$ is that the latter may contain infinite series while the former may not. Now we confirm that, indeed, $\lim F = \mbb{Z}[[x]]$, along with natural projections $
		\pi_n\colon\mbb{Z}[[x]]\to\Zpoly{n}
	$.

	In fact, the $f\colon R\to \lim F$ in \eqref{eq:polynomial_ring_tower} can be explicitly written down as:
	\begin{equation}
	\begin{aligned}
		f &= \eta_1 + x\,\pqty{
			\dv{x}\,\eta_2
		} + x^2\,\pqty{
			\frac{1}{2}\,\dv{x}\,\eta_3
		} + \cdots
		= \sum_{m=0}^\infty x^m\,\Bqty{
			\frac{1}{m!}\,\dv[m]{x}\,\eta_{m+1}
		} \\
		&\sim \pqty\big{
			\eta_{1,0},\,
			\eta_{2,1},\,
			\cdots,\,
			\eta_{n,n-1},\,
			\cdots
		}
	\end{aligned}
	\end{equation}
	Here $\frac{1}{m!}\,\dv[m]{x}$ is used to extract the $a_{n-1}$ coefficient in $\Zpoly{n}$; this is the last component of $\eta_n$, denoted by $\eta_{n,n-1}$. Any other choice of $f$ will break communitativity of \eqref{eq:polynomial_ring_tower}, hence $f$ is fixed uniquely by $\mbb{Z}[[x]]$ and $\pi_n$'s. Therefore, $\lim F = \mbb{Z}[[x]]$. \qed

	On the other hand, $\colim F$ is the ``largest'' object that any map \textit{out of} $\Zpoly{n}$ must ``passes through''. Also, this should hold for all $n\in\mbb{Z}_+$. Naturally, projections $
		\sigma_n\colon\Zpoly{n}\to\mbb{Z}
	$ satisfy the above requirements; we have:
	\begin{gather}
		\sigma_n\colon\ x\longmapsto 0,\quad
			\sum_{m=0}^{n-1} a_m x^m
			\longmapsto a_0,\\
		\sigma_{n+1}
		= p_1\circ p_2\circ\cdots p_n
	\end{gather}
	$g\colon \mbb{Z}\to R$ in \eqref{eq:polynomial_ring_tower} is fixed uniquely for such choice of $\sigma_n$; in fact, descend along the $p_n$ tower in \eqref{eq:polynomial_ring_tower}, and we have: $
		\tau_{n+1}
		= \tau_n\circ p_n
		= \tau_{n-1}\circ p_{n-1}\circ p_n
		= \cdots
		= \tau_1\circ p_1\circ p_2\circ\cdots p_n
		= \tau_1\circ \sigma_{n+1},
		\ \forall\ n\in\mbb{Z}_+
	$, hence $\exists\,!\ g = \tau_1$. Therefore, $\colim F = \mbb{Z}$. \qedfull

	\item \textbf{Example of push-out in \Cat{Groupoid}:}
	\begin{equation}
	\begin{tikzcd}[row sep=8.5ex,column sep=6.5ex]
		\{0,1\}
			\arrow[r,"f_1"]
			\arrow[d,"f_2",swap] &
		\bullet
			\arrow[d,"\tau_1"]
			\arrow[ddr,bend left,"\eta_1"] &[1ex] \\
		\{ 0\leftrightarrow 1 \}
			\arrow[r,"\tau_2"]
			\arrow[drr,bend right,"\eta_2"] &
		P
			\arrow[dr,dotted
				,"\!\!\exists!\,g" description]
		& \\[-2.5ex]
		&& Q
	\end{tikzcd}
	\end{equation}

	Following the same observation as before, the push-out $P$ is the ``largest'' object that any map out of $\bullet$ and $\{0\leftrightarrow  1\}$ must pass through. By such universal property, $P$ can be no larger than the coproduct: $
		\{\bullet\} \coprod \{0\leftrightarrow 1\}
	$. However, we should also consider the equivalence imposed by: 
	\begin{equation}
		\bullet
		\xlongleftarrow{\ f_1\,} \{0,1\}
		\xlongrightarrow{\ f_2\ }
		\{0\leftrightarrow 1\}
	\end{equation}
	Therefore, we simply have $P = \bullet$, with $\tau_{1,2}$ the natural projection. This can be verified with ease: we have $g = \eta_1$. It is unique since its image is a single point (with identity map to itself) $\star \in Q$, and the point $\star$ is fixed by commutativity. \qedfull

	\item \textbf{Product and coproduct in \Cat{Ab}:}

	For $G_\alpha\in\Cat{Ab}\subset\Cat{Group}$, note that we have:
	\begin{equation}
		\mop{Free}:
			\Cat{Set}
			\xrightleftharpoons{\quad}
			\Cat{Group}
		:\mop{Forget}
	\label{eq:free_forget_adjoint}
	\end{equation}
	Therefore, for $F$: some diagram in \Cat{Group}, $
		\boxed{\lim\,(\mop{Forget}\circ F)
		= \mop{Forget}\circ\lim F}
	$ if $\lim F$ exists.

	By definition, the product $\prod_\alpha G_\alpha\in \Cat{Group}$ is a limit, hence it is idential (as in $\Cat{Set}\mspace{.2mu}$) to the \textit{direct product}, with additional entry-wise group multiplication. Same applies for the full subcategory: abelian group $\Cat{Ab}\subset\Cat{Group}$.

	On the other hand, the disjoint union of $G_\alpha$'s as sets will not necessary be a group, the identities $\idty_\alpha\in G_\alpha$ must be glued together to produce a group structure. Furthermore, free-forgetful adjunction \eqref{eq:free_forget_adjoint} implies that for $F'$: some diagram in \Cat{Set},
	\begin{equation}
		\colim\,(\mop{Free}\circ F')
		= \mop{Free}\circ\colim F',\quad
	\end{equation}
	Whenever $\colim F'$ exists; in our case, $\colim F'$ is the disjoint union of sets: $
		\coprod_\alpha \mop{Forget}(G_\alpha)
	$. Therefore, we should construct a free object in \Cat{Ab}.

	Here we restrict our discussion to \Cat{Ab}, since the coproduct in $\Cat{Ab}$ is \textit{not} the same as in $\Cat{Group}$ --- the free product of abelian group is not necessary abelian. Hence, the coproduct in \Cat{Ab} shall be:
	\begin{equation}
		\coprod_\alpha G_\alpha
		= \bigoplus_\alpha G_\alpha,\quad
		i_\alpha\colon\, G_\alpha
			\ \xhookrightarrow{\quad}\ %
			\bigoplus_\alpha G_\alpha
	\end{equation}
	As a set, this is precisely the disjoint union with identities $0_\alpha \in G_\alpha \subset \Cat{Ab}$ glued together.

	It is then straight-forward to verify its universal property: for $f_\alpha\colon G_\alpha\to H$,
	\begin{equation}
		\exists\,!\ f\colon\,%
			\bigoplus_\alpha G_\alpha
			\ \longto\ H,\quad
		(g_\alpha)_\alpha
			\ \longmapsto\ %
			\sum_\alpha f_\alpha(g_\alpha)
	\end{equation}
	This is compatible with the abelian group multiplication. Note that for the summation to be well-defined, the coproduct must only contain finitely many components; otherwise it is identical to the product in \Cat{Ab}. 
	\qedfull

	\item \textbf{Composition of pull-backs:}
	\begin{equation}
	\begin{tikzcd}[row sep=7.5ex,column sep=7.5ex]
		Q
			\arrow[ddr,bend right,"g_1",swap]
			\arrow[dr,"f_1"]
			\arrow[drr,"f_2",bend left=12]
			\arrow[drrr,"f_3",bend left=20]
		&[1ex] && \\[-1ex]
		& A_1
			\arrow[rr,"\phi'" description
				,bend left,swap
				,crossing over]
			\arrow[r,dashed]
			\arrow[d] &
		A_2
			\arrow[d]
			\arrow[r,"\phi"] &
		A_3
			\arrow[d] \\
		& B_1
			\arrow[rr,"\psi'" description
				,bend right=25,swap]
			\arrow[r,"\psi",swap,dashed] &
		B_2
			\arrow[from=uull,"g_2",swap
				,bend right=20
				,crossing over]
			\arrow[r] &
		B_3
	\end{tikzcd}
	\end{equation}

	\begin{enumerate}
	\item If $A_1$ is the pull-back of $A_2$ and $A_2$ is the pull-back of $A_3$, then given $Q$ with $f_3,g_1$, we have $g_2 = \psi\circ g_1$, and $f_2$ is fixed uniquely by universal property of $A_2$, while $f_1$ is fixed uniquely by universal property of $A_1$. Hence, $A_1$ is the pull-back of $A_3$.

	\item If $A_1,A_2$ are pull-backs of $A_3$, then given $Q$ with $f_2,g_1$, we have $f_3 = \phi\circ f_2$, and $f_2$ is fixed uniquely by universal property of $A_2$, and $f_1$ is fixed uniquely by universal property of $A_1$. Hence, $A_1$ is the pull-back of $A_3$.

	\legacyReference

	\begin{enumerate}
	\item For $F_i \to E_i \xrightarrow{p_i} B$: coverings in $\mop{Cov}_0(B)$ with $E_i$: connected and $B$: path connected and locally path connected, the following diagram commutes:
	\begin{center}
	\begin{tikzcd}[row sep=5ex,column sep=3em]
		E_1
			\arrow{rr}{f}
			\arrow[swap]{dr}{p_1} & &
		E_2
			\arrow{dl}{p_2} \\
		& B &
	\end{tikzcd}
	\quad
	$\begin{gathered}
%		p_1 = p_2 \circ f,\\
		e_2 = f(e_1),\\
		b = p_1(e_1) = p_2(e_2),
	\end{gathered}$
	\end{center}
	To show that $f$ is itself a covering, we need only verify that $f$ is locally trivial with some discrete fiber $F$. 
	In fact, given any $e_2\in E_2$ and $b = p_2(e_2)$, there exists some neighborhood $U\subset B$ that the following diagram holds (by restriction):
	\begin{center}
	\begin{tikzcd}[row sep=5ex,column sep=2.5em]
		U\times F_1
			\arrow{rr}{f}
			\arrow[swap]{dr}{p_1} & &
		U\times F_2
			\arrow{dl}{p_2} \\
		& U &
	\end{tikzcd}
	\quad
	$\begin{aligned}
		e_1 &= \pqty\big{b,k_1},\\
		e_2 &= \pqty\big{b,k_2(b,k_1)},
	\end{aligned}
	\quad k_i \in F_i
	$
	\end{center}

>>>>

\newcommand{\hwNumber}{2}

@@

@@

	\begin{enumerate}
	\item \textbf{Morphism between coverings is covering:}

	For $F_i \to E_i \xrightarrow{p_i} B$: coverings in $\mop{Cov}_0(B)$ with $E_i$: connected and $B$: path connected and locally path connected, the following diagram commutes:
	\begin{center}
	\begin{tikzcd}[row sep=5ex,column sep=3em]
		E_1
			\arrow{rr}{f}
			\arrow[swap]{dr}{p_1} & &
		E_2
			\arrow{dl}{p_2} \\
		& B &
	\end{tikzcd}
	\quad
	$\begin{gathered}
%		p_1 = p_2 \circ f,\\
		e_2 = f(e_1),\\
		b = p_1(e_1) = p_2(e_2),
	\end{gathered}$
	\end{center}
	To show that $f$ is itself a covering, we need only verify that $f$ is locally trivial with some discrete fiber $F$. 
	In fact, given any $e_2\in E_2$ and $b = p_2(e_2)$, there exists some neighborhood $U\subset B$ that the following diagram holds (by restriction):
	\begin{center}
	\begin{tikzcd}[row sep=5ex,column sep=2.5em]
		U\times F_1
			\arrow{rr}{f}
			\arrow[swap]{dr}{p_1} & &
		U\times F_2
			\arrow{dl}{p_2} \\
		& U &
	\end{tikzcd}
	\quad
	$\begin{aligned}
		e_1 &= \pqty\big{b,k_1},\\
		e_2 &= \pqty\big{b,k_2(b,k_1)},
	\end{aligned}
	\quad k_i \in F_i
	$
	\end{center}
